% or a part of intro?
\section*{Literature Review} 
One of the key considerations in automating Slovenian traffic news is the availability and adaptability of LLMs for multilingual text generation. Studies have shown that most state-of-the-art models, including OpenAI’s GPT series \cite{brown2020language}, BLOOM \cite{scao2022bloom}, and mT5 \cite{xue2021mt5}, demonstrate strong multilingual capabilities. However, Slovenian, being a low-resource language, remains under-represented in large-scale training corpora \cite{ulcar2021sloberta}. Locally fine-tuned models such as SloT5 \cite{ulcar2022slot5} have emerged to address this gap, showing promise in domain-specific Slovenian text generation.

A crucial challenge in this domain is balancing between **fine-tuning** and **prompt engineering**. Fine-tuning LLMs on domain-specific text can improve accuracy but requires computational resources and well-annotated data \cite{zhang2022fine}. Prompt engineering, on the other hand, provides a lighter-weight alternative by designing effective input prompts to guide the model’s response \cite{reynolds2021prompt}. For structured and time-sensitive content like traffic news, a hybrid approach may be necessary, combining a base model with well-optimized prompting techniques.

Parthasarathy et al. propose and describe a seven stage pipeline for fine-tuning an LLM that covers everything from data preparation to model evaluation \cite{ultimate2024}.
It also provides an overview of (parameter-efficient) fine-tuning techniques, enabling the selection of the optimal approach based on specific performance objectives and time efficiency requirements.

Additionally, context-aware traffic reporting can benefit from external data sources, such as live weather updates, public holiday schedules, and road congestion analytics.
Studies have indicated that integrating real-time sensor data and probabilistic event modelling improves the predictive accuracy of automated reports \cite{schelter2019automated}.
One option to explore is to include such data in the prompt at runtime, resembling a retrieval-augmented generation (RAG) chain approach, only with live data instead of extensive documents.

Based on these insights, the following sections analyse the provided structured dataset of Slovenian traffic news, examining patterns in report generation and urgency to inform an AI-based automation approach.
